---
title: "Data_Analysis"
author: "Terry Slenn"
date: "November 23, 2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(gsheet)
library(ggplot2)
library(ggmap)
```


```{r gsheets}
Prices <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1LC1dk9HlzCEqVPN2AGI6trGaXpwlP8wtyYheK6k700A/edit?usp=sharing')

Prices <- Prices %>% 
  separate(LAT_LONG, into = c("LAT", "LONG"), sep = ", ") %>% 
  mutate(LAT = as.numeric(LAT), LONG = as.numeric(LONG))

Stores <- Prices %>% 
  distinct(ADDRESS, LAT, LONG, STORE, STATE)

Prices %>% 
  distinct(`ZIP CODE`, STATE) %>% 
  arrange(STATE)
```


```{r ggmap}
DC_map <- get_map(location = c(lat = 38.937495, lon =-77.088846), zoom = 11)
ggmap(DC_map) +
  geom_point(data = Stores, mapping = aes(x = LONG, y = LAT, fill = STATE),
             pch = 21, size = 2.5) +
  theme_void() +
  theme(legend.position = "none")+
  scale_fill_viridis_d()
```

## Data Cleaning

```{r data cleaning}
OJ <- Prices %>% 
  filter(PRODUCT == "Orange Juice") %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE)*64) ## Standardize to 64 oz

Milk <- Prices %>% 
  filter(PRODUCT == "Milk") %>% 
  mutate(PRICE_STD = PRICE) ## SHOULD all be the same

Eggs <- Prices %>% 
  filter(PRODUCT == "Eggs") %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE)*12)

Beef <- Prices %>% 
  filter(PRODUCT == "Ground Beef") %>% 
  mutate(PRICE_STD = PRICE)

Potatoes <- Prices %>% 
  filter(str_detect(PRODUCT, "ota")) %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE),
         PRICE_STD = coalesce(PRICE_STD, PRICE))

Cola <- Prices %>% 
  filter(PRODUCT == "Cola") %>% 
  mutate(PRICE_STD = PRICE)

Tortilla <- Prices %>% 
  filter(str_detect(PRODUCT, "illa")) %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE)*12)

Prices <- bind_rows(OJ, Milk, Eggs, Beef, Potatoes, Cola, Tortilla) %>% 
  group_by(ADDRESS, LAT, LONG, `ZIP CODE`, STATE, STORE) %>% 
  summarize(All_Goods = sum(PRICE_STD),
            Luxury = sum(PRICE_STD[PRODUCT %in% c("Cola", "Tortilla Chips")]),
            Essential = All_Goods - Luxury)

```

## Exploratory Analysis

```{r pricemap}
Store_Price <- Stores %>% 
  left_join(Prices)

ggmap(DC_map) +
  geom_point(data = Store_Price, mapping = aes(x = LONG, y = LAT, fill = All_Goods),
             pch = 21, size = 2.5) +
  theme_void() +
  scale_fill_gradient(low = "green", high = "red")
  #scale_fill_viridis_d()
```

```{r box plots}
Prices %>% 
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>% 
  ggplot(aes(x = STATE, y = Total_Price, color = STATE)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~ Basket, scales = "free_y") +
  ggtitle("Basket of Goods Price: Region Comparison") +
  theme(legend.position = "none")

Prices %>% 
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>%
  ggplot(aes(x = STORE, y = Total_Price, color = STORE)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~ Basket, scales = "free_y", nrow = 2) +
  #coord_flip() +
  ggtitle("Basket of Goods Price: Store Comparison") +
  theme(legend.position = "none")
```

## Further work needed

Need to perform formal statistics tests between groups and figure out how cluster sampling impacts that.  

Need to pull in demographic info to evaluate any trends that may exist with demographics and income.



## Cluster Analysis

Primary Sampling Unit (PSU): Zip Code  

Secondary Sampling Unit (SSU): Grocery Store  

Estimated average price for each zip code:  

$\hat{\bar{y}}_i = \frac{\sum{t_i}}{\sum{m_i}}$  

$Var(\hat{\bar{y}}_i) = Var(\frac{\sum{t_i}}{\sum{m_i}})$  

$Var(\hat{\bar{y}}_i) = \frac{1}{(\sum{m_i})^2}  Var(\sum{t_i})$  

$Var(\hat{\bar{y}}_i) = \frac{1}{(\sum{m_i})^2}  Var(\sum{t_i})$



```{r cluster}
DC <- Prices %>%  
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>% 
  group_by(Basket, STATE) %>% 
  filter(STATE == "DC") %>% 
  summarize(Mean_hat = mean(Total_Price),
            Var_hat = sum((Total_Price - Mean_hat)^2)/n(),
            M = n())

Cluster_est <- Prices %>% 
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>% 
  filter(STATE != "DC") %>% 
  group_by(STATE, `ZIP CODE`, Basket) %>% 
  summarize(Mean = mean(Total_Price),
            Var = sum((Total_Price - Mean)^2)/n(),
            m = n()) %>% 
  group_by(STATE, Basket) %>% 
  summarize(Mean_hat = sum(m*Mean)/sum(m),
            Var_hat = sum(m*Var)/sum(m),
            M = sum(m)) %>% 
  rbind(DC)
```


```{r cluster_graph}
Cluster_est %>% 
  ggplot(aes(x = STATE, y = Mean_hat, color = STATE)) +
  geom_errorbar(aes(ymin = Mean_hat - qt(.975,M-1), ymax = Mean_hat + qt(.975,M-1))) +
  facet_wrap(~ Basket, scales = "free_y") +
  ggtitle("Basket of Goods Price: Region Comparison") +
  theme(legend.position = "none")
```


## Demographics

```{r demo_data}

Demographics <- gsheet2tbl('https://docs.google.com/spreadsheets/d/10TozoY2R9g64xsm9vJxvWwuuV4EX-CZic7RV2XhKCh4/edit?usp=sharing') %>% 
  select(-Area)

Prices_demo <- left_join(Prices, Demographics)

Prices_demo %>%
  ungroup() %>% 
  select(-ADDRESS, -LAT, -LONG, -`ZIP CODE`, -STORE, -STATE, - Luxury, -Essential) %>% 
  plot()

Prices_demo %>% 
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>%
  filter(Basket == "Luxury") %>% 
  ggplot(aes(y = Total_Price, x = `poverty rate`, color = STATE)) +
  geom_point() #+
  #facet_wrap(~Basket, scale = "free_y")
  #geom_smooth(method = "lm")

lm(Luxury ~ `poverty rate`, data = Prices_demo) %>% 
  summary()
```


```{r analysis}
Prices_demo %>% 
  ggplot(aes(x=`Population Density`, y = All_Goods)) +
  geom_point(aes(color = STATE)) +
  geom_smooth(method = "lm", se = FALSE)
lm(All_Goods ~ `Population Density`, data = Prices_demo) %>% 
  summary()

Prices_demo %>% 
  ggplot(aes(x=`Population Density`, y = Luxury)) +
  geom_point(aes(color = STATE)) +
  geom_smooth(method = "lm", se = FALSE)
lm(Luxury ~ `Population Density`, data = Prices_demo) %>% 
  summary()

lm(Luxury ~ `Population Density` + `poverty rate`, data = Prices_demo) %>% 
  summary()

lm(All_Goods ~ `poverty rate` + `median income`, data = Prices_demo) %>% 
  summary()
```
