---
title: "Data_Analysis"
author: "Terry Slenn"
date: "November 23, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(gsheet)
library(ggplot2)
library(ggmap)
```


```{r gsheets}
Prices <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1LC1dk9HlzCEqVPN2AGI6trGaXpwlP8wtyYheK6k700A/edit?usp=sharing')

Prices <- Prices %>% 
  separate(LAT_LONG, into = c("LAT", "LONG"), sep = ", ") %>% 
  mutate(LAT = as.numeric(LAT), LONG = as.numeric(LONG))

Stores <- Prices %>% 
  distinct(ADDRESS, LAT, LONG, STORE, STATE)

Prices %>% 
  distinct(`ZIP CODE`, STATE)
```


```{r ggmap}
DC_map <- get_map(location = c(lat = 38.937495, lon =-77.088846), zoom = 11)
ggmap(DC_map) +
  geom_point(data = Stores, mapping = aes(x = LONG, y = LAT, fill = STATE),
             pch = 21, size = 2) +
  theme_void() +
  theme(legend.position = "none")+
  scale_fill_viridis_d()
```

## Data Cleaning

```{r data cleaning}
OJ <- Prices %>% 
  filter(PRODUCT == "Orange Juice") %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE)*64) ## Standardize to 64 oz

Milk <- Prices %>% 
  filter(PRODUCT == "Milk") %>% 
  mutate(PRICE_STD = PRICE) ## SHOULD all be the same

Eggs <- Prices %>% 
  filter(PRODUCT == "Eggs") %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE)*64)

Beef <- Prices %>% 
  filter(PRODUCT == "Ground Beef") %>% 
  mutate(PRICE_STD = PRICE)

Potatoes <- Prices %>% 
  filter(str_detect(PRODUCT, "ota")) %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE),
         PRICE_STD = coalesce(PRICE_STD, PRICE))

Cola <- Prices %>% 
  filter(PRODUCT == "Cola") %>% 
  mutate(PRICE_STD = PRICE)

Tortilla <- Prices %>% 
  filter(str_detect(PRODUCT, "illa")) %>% 
  mutate(PRICE_STD = PRICE/parse_number(SIZE)*12)

Prices <- bind_rows(OJ, Milk, Eggs, Beef, Potatoes, Cola, Tortilla) %>% 
  group_by(ADDRESS, LAT, LONG, `ZIP CODE`, STATE, STORE) %>% 
  summarize(All_Goods = sum(PRICE_STD),
            Luxury = sum(PRICE_STD[PRODUCT %in% c("Cola", "Tortilla Chips")]),
            Essential = All_Goods - Luxury)

```

## Exploratory Analysis

```{r pricemap}
Store_Price <- Stores %>% 
  left_join(Prices)

ggmap(DC_map) +
  geom_point(data = Store_Price, mapping = aes(x = LONG, y = LAT, fill = All_Goods),
             pch = 21, size = 2) +
  theme_void() +
  scale_fill_gradient(low = "green", high = "red")
  #scale_fill_viridis_d()
```

```{r box plots}
Prices %>% 
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>% 
  ggplot(aes(x = STATE, y = Total_Price, color = STATE)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~ Basket, scales = "free_y") +
  ggtitle("Basket of Goods Price: Region Comparison") +
  theme(legend.position = "none")

Prices %>% 
  gather(key = "Basket", value = "Total_Price", All_Goods, Essential, Luxury) %>%
  ggplot(aes(x = STORE, y = Total_Price, color = STORE)) +
  geom_boxplot() +
  geom_jitter() +
  facet_wrap(~ Basket, scales = "free_y", nrow = 2) +
  #coord_flip() +
  ggtitle("Basket of Goods Price: Store Comparison") +
  theme(legend.position = "none")
```

## Further work needed

Need to perform formal statistics tests between groups and figure out how cluster sampling impacts that.  

Need to pull in demographic info to evaluate any trends that may exist with demographics and income.







**Problems:**
  Milk was supposed to be gallon price.  Arlington data is throwing us way off.
  8 lb bag of potatoes is also messing up data



